---
title: "Analyzing"
author: "Mark Barrett"
format: html
editor_options: 
  chunk_output_type: console
---

# Plan

Analyze the differences in travel time to Los Angeles 2028 Olympic venues between driving and transit.
Collect tract-level travel time data to each venue using both transit and driving modes.
To do this, I will find the centroid of each tract in the Los Angeles area and then use the TravelTime API to collect travel time data to each venue.

- 30 mins, 1hr, and 1.5hrs
- via car and via public transport
- at 8am and 11am on a Weekday

```{r setup}
library(here)
library(tidyverse)
library(sf)
library(traveltimeR)
library(jsonlite)

# Load custom functions
source(here("R", "fct-traveltime.R"))
```

# Data

## Census Tracts

Downloaded from nhgis.com

```{r}
# Census tract shapefiles
nhgis_shp_file = here("data", "nhgis", "US_tract_2020.shp")

# Create list of tracts to exclude
exclude_tracts = tribble(
  ~TRACTCE, ~include,
  "599100", F, # Catalina Island
  "599000", F, # Catalina Island
  "930400", F,  # Angeles Forest
)

# Load the shapefile
nhgis_shp = read_sf(nhgis_shp_file) |> 
  # Filter to Los Angeles, CA
  filter(STATEFP == "06", COUNTYFP == "037") |>
  # Reproject CRS to WGS84
  st_transform(crs = "EPSG:4326") |> 
  # Create an opinionated subset of LA tracts
  # anti_join(exclude_tracts, by = "TRACTCE") |>
  # Remove tracts where the centroid is north of 34.35
  # filter(st_coordinates(st_centroid(geometry))[,2] < 34.35) |>
  select(GISJOIN, TRACTCE, GEOID, NAME, NAMELSAD, geometry) |> 
  left_join(exclude_tracts, by = "TRACTCE") |> 
  mutate(include = case_when(!is.na(include) ~ include,
                             st_coordinates(st_centroid(geometry))[,2] > 34.35 ~ F,
                             .default = T))

glimpse(nhgis_shp)

```

Check how this looks on a map

```{r}
ggplot(nhgis_shp) +
  geom_sf(aes(fill = include), colour=NA) +
  coord_sf() +
  labs(title = "Census Tracts in Los Angeles County",
       subtitle = "Filtered to exclude Catalina Island, Angeles Forest, and Antelope Valley")+
  theme_void()
```

Remove these tracts

```{r}
nhgis_la = nhgis_shp |> 
  filter(include) |> 
  select(-include)
```


### Find centroids of census tracts

```{r}
nhgis_centroids = nhgis_la |> 
  # Find centroids of each tract
  mutate(centroid = st_centroid(geometry))

ggplot(nhgis_centroids) +
  geom_sf(aes(geometry = geometry), fill = NA) +
  geom_sf(aes(geometry = centroid), colour = "red")+
  coord_sf() +
  labs(title = "Census Tracts centroids")+
  theme_void()
```

Okay so there are all our centroids, how many do we have in total?

```{r}
nrow(nhgis_centroids)
```

Next, let's pull in the venues data

## Venues

```{r}
venues = read_rds(here("data", "venues_complete.rds"))

glimpse(venues)
```

## Travel Times

```{r}
travel_times = read_rds(here("data", "travel_times_2025-07-24.rds"))

glimpse(travel_times)
```

```{r}
count(travel_times, venue_id, sort = T)
```

For some reason, the `venue_id`s have a number between 1 and 2 appended to them.

```{r}
travel_times |> 
  filter(str_detect(venue_id, "LA Live")) |> View()
```

```{r}
travel_times_clean = travel_times |> 
  mutate(venue = str_remove_all(venue_id, "[12]\\s\\|\\s\\w+"), .after = "id") 

travel_times_pivot = travel_times_clean |> 
  pivot_wider(id_cols = c(id, venue), names_from = transportation, names_glue = "{transportation}_{.value}", values_from = c(time_mins, distance_mi)) |> 
  rename_with(.fn = \(x) str_remove_all(x, "_mins"))
```

What are the average travel times for each venue?

```{r}
travel_times_pivot |> 
  group_by(venue) |> 
  summarise(across(contains("_time"), list(min = min, mean = mean, median = median, max = max), na.rm = T))
```

```{r}
travel_times_clean |> 
  group_by(venue, transportation) |> 
  summarise(across(time_mins, list(min = min, mean = mean, median = median, max = max), na.rm = T))
```


```{r}
ggplot(travel_times_clean, aes(x = time_mins, colour = transportation, fill = transportation))+
    geom_density()+
    facet_grid(rows = vars(venue), scales = "free_y", switch = "y")+
  theme_minimal()+
  theme(strip.text.y.left = element_text(angle = 0), strip.placement = "outside")

```

What is the average difference in travel time between driving and transit?

```{r}
travel_times_pivot |> 
  mutate(time_diff = (public_transport_time - driving_time*1.1), .after = "public_transport_time") |> 
  group_by(venue) |> 
  summarise(across(time_diff, list(min = min, mean = mean, median = median, max = max), na.rm = T))
  
```

