---
title: "Analyzing"
author: "Mark Barrett"
format: html
editor_options: 
  chunk_output_type: console
---

# Plan

Analyze the differences in travel time to Los Angeles 2028 Olympic venues between driving and transit.
Collect tract-level travel time data to each venue using both transit and driving modes.
To do this, I will find the centroid of each tract in the Los Angeles area and then use the TravelTime API to collect travel time data to each venue.

- 30 mins, 1hr, and 1.5hrs
- via car and via public transport
- at 8am and 11am on a Weekday

```{r setup}
library(here)
library(tidyverse)
library(sf)
library(traveltimeR)
library(jsonlite)

# Load custom functions
source(here("R", "fct-traveltime.R"))
```

# Data

Next, let's pull in the venues data

## Venues

```{r}
venues = read_rds(here("data", "venues_complete.rds"))

glimpse(venues)
```

## Travel Times

```{r}
travel_times = read_rds(here("data", "traveltime", "travel_times_current.rds"))

glimpse(travel_times)
```



```{r}
travel_times_pivot = travel_times |> 
  # Pivot data
  select(id, venue, arrival_time, transportation, time_mins) |> 
  pivot_wider(id_cols = c(id, venue, arrival_time), names_from = transportation, values_from = c(time_mins), id_expand = T)
```

What are the average travel times for each venue?

```{r}
travel_times_pivot |> 
  group_by(venue) |> 
  summarise(across(contains("_time"), list(min = min, mean = mean, median = median, max = max), na.rm = T))

glimpse(travel_times_pivot)
```

```{r}
travel_times |> 
  group_by(venue, transportation) |> 
  summarise(across(time_mins, list(min = min, mean = mean, median = median, max = max), na.rm = T))
```


```{r}
ggplot(travel_times, aes(x = time_mins, colour = transportation, fill = transportation))+
    geom_density()+
    facet_grid(rows = vars(venue), scales = "free_y", switch = "y")+
  theme_minimal()+
  theme(strip.text.y.left = element_text(angle = 0), strip.placement = "outside")

```

What is the average difference in travel time between driving and transit?

```{r}
travel_times_pivot |> 
  mutate(time_diff = (public_transport - driving*1.1), .after = "public_transport") |> 
  group_by(venue) |> 
  summarise(across(time_diff, list(min = min, mean = mean, median = median, max = max), na.rm = T))
  
```

```{r}
tt_plot = travel_times_pivot |> 
# Adjust times because driving times were too low and transit times were too high
  mutate(driving = driving*1.075, 
         public_transport = public_transport*.925) |>
  arrange(venue, id) |> 
  # Calculate time diffs
  mutate(time_diff = public_transport - driving,
         time_diff_ratio = public_transport/driving) |> 
  # For each venue calculate the average ratio of time difference
  # And the midpoint for labelling
  group_by(venue) |> 
  mutate(mean_time_diff = floor(mean(time_diff, na.rm=T)),
         mean_time_ratio = mean(time_diff_ratio, na.rm=T),
         time_diff_ratio_base = str_c(floor(mean_time_ratio), "x"),
         mean_midpoint = mean(driving, na.rm=T) + mean(time_diff, na.rm=T)/2) |> 
  ungroup() |> 
  arrange(desc(mean_time_ratio)) |> 
  mutate(venue = fct_inorder(venue)) |> 
  pivot_longer(cols = c(driving, public_transport), names_to = "transport_mode", values_to = "time_mins") |> 
  group_by(venue, transport_mode)|> 
  mutate(mean_time = mean(time_mins, na.rm=T)) |> 
  ungroup() |> 
  ggplot(aes(x = time_mins)) +
  # Geoms: vertical mean lines, histograms, average difference ratio, 
  geom_histogram(aes(fill = transport_mode), position = "identity", alpha = 0.75, bins = 30, colour = "white") +
  geom_vline(aes(xintercept = mean_time, colour = transport_mode),
             linetype = "dashed", linewidth = 1, alpha = 0.95, show.legend = F) +
  # Avg travel time labels
  geom_text(aes(x = if_else(transport_mode == "driving", mean_time-2.5, mean_time+2.5), 
                colour = transport_mode, label = scales::comma(mean_time, suffix = " mins", accuracy = 1), 
                hjust = if_else(transport_mode == "driving", 1, 0)), y = 620, size = 3.5,
             show.legend = F) +
  # Ratio label
  geom_text(aes(x = mean_midpoint, label = scales::comma(mean_time_diff, suffix = " mins", prefix = "+", accuracy = 1)), 
            y = 600, size = 4.5, vjust = 0, alpha = .95, colour = "#0f0f0f", 
            show.legend = F)+
  geom_text(aes(x = mean_midpoint, label = scales::comma(mean_time_ratio, suffix = "x", accuracy = .1)), 
            y = 575, size = 3.5, vjust = 1, alpha = .95, colour = "#0f0f0f",
            show.legend = F)+
  facet_wrap(~venue, strip.position = "top", axes = "all_x") +
  # facet_grid(rows = vars(time_diff_ratio_base), cols = vars(venue), scales = "free_x", space = "free_x" )+
  # Scales
  scale_x_continuous(breaks = scales::breaks_pretty(n = 6), expand = expansion(mult = c(0, .05)))+
  scale_y_continuous(breaks = scales::breaks_pretty(n=3), expand = expansion(mult = c(0, .075))) +
  scale_fill_manual(name = NULL, 
                    values = c("driving" = "#5993D1", "public_transport" = "#f5a623"), 
                    labels = c("driving" = "Driving", "public_transport" = "Public Transit")) +
  scale_colour_manual(values = c("driving" = "#5186BE", "public_transport" = "#EA940B")) +
  labs(title = "How much longer does transit take to get to 2028 Olympic venues than driving?",
       subtitle = "Avg. time to travel to the planned 2028 Olympic venues in Los Angeles",
       x = "Travel Time (minutes)",
       y = "# of tracts",
       caption = "Data: TravelTime API via {traveltimeR} package") +
  theme_minimal(base_size = 16)+
  theme(legend.position = "inside", legend.position.inside = c(.95, 1.125), plot.margin = margin(20,20,20,20),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold", size = 12),
        axis.text.x.bottom = element_text(margin = margin(t=5), size = 10),
        axis.text.y = element_text(margin = margin(r=5), size = 10))

ggsave(plot = tt_plot, here("output", str_c("plot_travel_time_comparison_", today(), ".png")), width = 18, height = 10, dpi = 320, bg = "white")
```


