---
title: "Analyzing"
author: "Mark Barrett"
format: html
editor_options: 
  chunk_output_type: console
---

# Plan

Analyze the differences in travel time to Los Angeles 2028 Olympic venues between driving and transit.
Collect tract-level travel time data to each venue using both transit and driving modes.
To do this, I will find the centroid of each tract in the Los Angeles area and then use the TravelTime API to collect travel time data to each venue.

- 30 mins, 1hr, and 1.5hrs
- via car and via public transport
- at 8am and 11am on a Weekday

```{r setup}
library(here)
library(tidyverse)
library(sf)
library(traveltimeR)
library(jsonlite)

# Load custom functions
source(here("R", "fct-traveltime.R"))
```

# Data

Next, let's pull in the venues data

## Venues

```{r}
venues = read_rds(here("data", "venues_complete.rds"))

glimpse(venues)
```

## Census Tracts


```{r}
# 2020 population by census tract from: https://www.nhgis.org/
nhgis_file = here("data", "nhgis", "nhgis0001_ds258_2020_tract.csv")

nhgis = read_csv(nhgis_file)

# Filter for LA County
nhgis_la = nhgis |> 
  filter(STATE == "California", COUNTY == "Los Angeles County") |> 
  mutate(AREALAND_SQKM = AREALAND/1e6,
         AREALAND_SQMI = AREALAND_SQKM/2.59,
         .after = "AREALAND") |> 
  select(GISJOIN, YEAR, TRACTA, POPULATION = U7H001, AREALAND_SQKM, AREALAND_SQMI)

# Census tract shapefiles
nhgis_shp_file = here("data", "nhgis", "US_tract_2020.shp")

# Reproject CRS to WGS84
nhgis_shp = st_read(nhgis_shp_file) |> 
  st_transform(crs = "EPSG:4326")

```


Create an opinionated subset of Los Angeles tracts that make up the area of LA where transportation to venues is relevant.

Exclude Catalina Island (tract 5991) and the area north of the San Gabriel Mountains (tract 9304).
Also exclude tracts that are north of the 34.35 latitude: functionally Santa Clarita and above.

```{r}
# Join tract population and GIS data
nhgis_la_shp = inner_join(nhgis_la, nhgis_shp, by = "GISJOIN")

# Create list of tracts to exclude
exclude_tracts = tribble(
  ~TRACTCE,
  "599100", # Catalina Island
  "599000", # Catalina Island
  "930400" # Angeles Forest
)

# Create an opinionated subset of LA tracts
nhgis_la_shp_subset = nhgis_la_shp |> 
  anti_join(exclude_tracts, by = "TRACTCE") |>
  # Remove tracts where the centroid is north of 34.35
  filter(st_coordinates(st_centroid(geometry))[,2] < 34.35) |>
  st_as_sf()

glimpse(nhgis_la_shp_subset)

# st_write(nhgis_la_shp_subset, here("data", "nhgis_la_subset.geojson.json"), driver = "GeoJSON")
```

## Travel Times

```{r}
travel_times = read_rds(here("data", "traveltime", "travel_times_current.rds"))

glimpse(travel_times)
```



```{r}
travel_times_pivot = travel_times |> 
  # Pivot data
  select(id, venue, arrival_time, transportation, time_mins) |> 
  pivot_wider(id_cols = c(id, venue, arrival_time), names_from = transportation, values_from = c(time_mins), id_expand = T)
```

What are the average travel times for each venue?

```{r}
travel_times_pivot |> 
  group_by(venue) |> 
  summarise(across(contains("_time"), list(min = min, mean = mean, median = median, max = max), na.rm = T))

glimpse(travel_times_pivot)
```

```{r}
travel_times |> 
  group_by(venue, transportation) |> 
  summarise(across(time_mins, list(min = min, mean = mean, median = median, max = max), na.rm = T))
```


```{r}
ggplot(travel_times, aes(x = time_mins, colour = transportation, fill = transportation))+
    geom_density()+
    facet_grid(rows = vars(venue), scales = "free_y", switch = "y")+
  theme_minimal()+
  theme(strip.text.y.left = element_text(angle = 0), strip.placement = "outside")

```

What is the average difference in travel time between driving and transit?

```{r}
travel_times_pivot |> 
  mutate(time_diff = (public_transport - driving*1.1), .after = "public_transport") |> 
  group_by(venue) |> 
  summarise(across(time_diff, list(min = min, mean = mean, median = median, max = max), na.rm = T))
  
```

```{r}
tt_plot = travel_times_pivot |> 
# Adjust times because driving times were too low and transit times were too high
  mutate(driving = driving*1.075, 
         public_transport = public_transport*.925) |>
  arrange(venue, id) |> 
  # Calculate time diffs
  mutate(time_diff = public_transport - driving,
         time_diff_ratio = public_transport/driving) |> 
  # For each venue calculate the average ratio of time difference
  # And the midpoint for labelling
  group_by(venue) |> 
  mutate(mean_time_diff = floor(mean(time_diff, na.rm=T)),
         mean_time_ratio = mean(time_diff_ratio, na.rm=T),
         time_diff_ratio_base = str_c(floor(mean_time_ratio), "x"),
         mean_midpoint = mean(driving, na.rm=T) + mean(time_diff, na.rm=T)/2) |> 
  ungroup() |> 
  arrange(desc(mean_time_ratio)) |> 
  mutate(venue = fct_inorder(venue)) |> 
  pivot_longer(cols = c(driving, public_transport), names_to = "transport_mode", values_to = "time_mins") |> 
  group_by(venue, transport_mode)|> 
  mutate(mean_time = mean(time_mins, na.rm=T)) |> 
  ungroup() |> 
  ggplot(aes(x = time_mins)) +
  # Geoms: vertical mean lines, histograms, average difference ratio, 
  geom_histogram(aes(fill = transport_mode), position = "identity", alpha = 0.75, bins = 30, colour = "white") +
  geom_vline(aes(xintercept = mean_time, colour = transport_mode),
             linetype = "dashed", linewidth = 1, alpha = 0.95, show.legend = F) +
  # Avg travel time labels
  geom_text(aes(x = if_else(transport_mode == "driving", mean_time-2.5, mean_time+2.5), 
                colour = transport_mode, label = scales::comma(mean_time, suffix = " mins", accuracy = 1), 
                hjust = if_else(transport_mode == "driving", 1, 0)), y = 620, size = 3.5,
             show.legend = F) +
  # Ratio label
  geom_text(aes(x = mean_midpoint, label = scales::comma(mean_time_diff, suffix = " mins", prefix = "+", accuracy = 1)), 
            y = 600, size = 4.5, vjust = 0, alpha = .95, colour = "#0f0f0f", 
            show.legend = F)+
  geom_text(aes(x = mean_midpoint, label = scales::comma(mean_time_ratio, suffix = "x", accuracy = .1)), 
            y = 575, size = 3.5, vjust = 1, alpha = .95, colour = "#0f0f0f",
            show.legend = F)+
  facet_wrap(~venue, strip.position = "top", axes = "all_x") +
  # facet_grid(rows = vars(time_diff_ratio_base), cols = vars(venue), scales = "free_x", space = "free_x" )+
  # Scales
  scale_x_continuous(breaks = scales::breaks_pretty(n = 6), expand = expansion(mult = c(0, .05)))+
  scale_y_continuous(breaks = scales::breaks_pretty(n=3), expand = expansion(mult = c(0, .075))) +
  scale_fill_manual(name = NULL, 
                    values = c("driving" = "#5993D1", "public_transport" = "#f5a623"), 
                    labels = c("driving" = "Driving", "public_transport" = "Public Transit")) +
  scale_colour_manual(values = c("driving" = "#5186BE", "public_transport" = "#EA940B")) +
  labs(title = "How much longer does transit take to get to 2028 Olympic venues than driving?",
       subtitle = "Avg. time to travel to the planned 2028 Olympic venues in Los Angeles",
       x = "Travel Time (minutes)",
       y = "# of tracts",
       caption = "Data: TravelTime API via {traveltimeR} package") +
  theme_minimal(base_size = 16)+
  theme(legend.position = "inside", legend.position.inside = c(.95, 1.125), plot.margin = margin(20,20,20,20),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        strip.text = element_text(face = "bold", size = 12),
        axis.text.x.bottom = element_text(margin = margin(t=5), size = 10),
        axis.text.y = element_text(margin = margin(r=5), size = 10))

ggsave(plot = tt_plot, here("output", str_c("plot_travel_time_comparison_", today(), ".png")), width = 18, height = 10, dpi = 320, bg = "white")
```

```{r}
tt_plot
```


For each tract, what is the average travel time for driving and transit?

```{r}
travel_times_summary_stats = travel_times_pivot |> 
  # filter(id == "101122") |>
  # arrange(public_transport)
  # mutate(public_transport = if_else(is.nan(public_transport), NA_real_, public_transport)) |> 
  group_by(id) |> 
  # summarise(across(c(driving, public_transport), \(x) mean(x, na.rm = T))) |> 
  summarise(across(c(driving, public_transport), list(mean = \(x) mean(x, na.rm = T),
                                                      median = \(x) median(x, na.rm = T),
                                                      min = \(x) min(x, na.rm = T),
                                                      max = \(x) max(x, na.rm = T)), .names = "{.fn}_{.col}"), .groups = "drop") |>
  pivot_longer(cols = !id, names_to = c("stat", "transportation"), names_sep = "(?<=m[a-z]{2,5})_", values_to = "time_mins")

nhgis_la_shp_subset |> 
  select(GISJOIN, TRACTA, geometry) |> 
  expand_grid(transportation = c("driving", "public_transport"), stat = c("mean", "median", "min", "max")) |> 
  left_join(travel_times_summary_stats, by = c("TRACTA" = "id", "transportation" = "transportation", "stat" = "stat")) |>
  ggplot(aes(fill = time_mins))+
  geom_sf(aes(geometry = geometry))+
  scale_fill_viridis_b(direction = -1, nice.breaks = T)+
  facet_grid(rows = vars(transportation), cols = vars(stat))+
  theme_void()
```

What does the map look like for one venue?

```{r}
one_venue_df = travel_times |> 
  filter(venue == sample(venue, 1)) |> 
  inner_join(venues, by = c("venue" = "venue_location"))

nhgis_la_shp_subset |> 
  select(GISJOIN, TRACTA, geometry) |> 
  expand_grid(transportation = c("driving", "public_transport")) |> 
  left_join(one_venue_df, by = c("TRACTA" = "id", "transportation" = "transportation")) |> 
  ggplot(aes(fill = time_mins))+
  geom_sf(aes(geometry = geometry))+
  geom_sf(aes(geometry = venue_geometry), size = 4)+
  geom_sf_text(aes(geometry = venue_geometry, label = venue), nudge_y = -0.05, size = 5, hjust = 1)+
  # scale_fill_viridis_b(direction = -1, option = "C", nice.breaks = T)+
  scale_fill_distiller(name = NULL, type = "div", palette = 8)+
  facet_wrap(~transportation)+
  theme_void()
```


