---
title: "Collecting Tract-level Travel Time Data"
author: "Mark Barrett"
format: html
---

Documentation:

- [Github](https://github.com/traveltime-dev/traveltime-sdk-r)
- [Routes](https://playground.traveltime.com/routes)

# Plan

Analyze the differences in travel time to Los Angeles 2028 Olympic venues between driving and transit.
Collect tract-level travel time data to each venue using both transit and driving modes.
To do this, I will find the centroid of each tract in the Los Angeles area and then use the TravelTime API to collect travel time data to each venue.

- 30 mins, 1hr, and 1.5hrs
- via car and via public transport
- at 8am and 11am on a Weekday

```{r setup}
library(here)
library(tidyverse)
library(sf)
library(traveltimeR)
library(jsonlite)

# Load custom functions
source(here("R", "fct-traveltime.R"))
```

# Data

## Census Tracts

Downloaded from nhgis.com

```{r}
# Census tract shapefiles
nhgis_shp_file = here("data", "nhgis", "US_tract_2020.shp")

# Create list of tracts to exclude
exclude_tracts = tribble(
  ~TRACTCE, ~include,
  "599100", F, # Catalina Island
  "599000", F, # Catalina Island
  "930400", F,  # Angeles Forest
)

# Load the shapefile
nhgis_shp = read_sf(nhgis_shp_file) |> 
  # Filter to Los Angeles, CA
  filter(STATEFP == "06", COUNTYFP == "037") |>
  # Reproject CRS to WGS84
  st_transform(crs = "EPSG:4326") |> 
  # Create an opinionated subset of LA tracts
  # anti_join(exclude_tracts, by = "TRACTCE") |>
  # Remove tracts where the centroid is north of 34.35
  # filter(st_coordinates(st_centroid(geometry))[,2] < 34.35) |>
  select(GISJOIN, TRACTCE, GEOID, NAME, NAMELSAD, geometry) |> 
  left_join(exclude_tracts, by = "TRACTCE") |> 
  mutate(include = case_when(!is.na(include) ~ include,
                             st_coordinates(st_centroid(geometry))[,2] > 34.35 ~ F,
                             .default = T))

glimpse(nhgis_shp)

```

Check how this looks on a map

```{r}
ggplot(nhgis_shp) +
  geom_sf(aes(fill = include), colour=NA) +
  coord_sf() +
  labs(title = "Census Tracts in Los Angeles County",
       subtitle = "Filtered to exclude Catalina Island, Angeles Forest, and Antelope Valley")+
  theme_void()
```

Remove these tracts

```{r}
nhgis_la = nhgis_shp |> 
  filter(include) |> 
  select(-include)
```


# Find centroids of census tracts

```{r}
nhgis_centroids = nhgis_la |> 
  # Find centroids of each tract
  mutate(centroid = st_centroid(geometry))

ggplot(nhgis_centroids) +
  geom_sf(aes(geometry = geometry), fill = NA) +
  geom_sf(aes(geometry = centroid), colour = "red")+
  coord_sf() +
  labs(title = "Census Tracts centroids")+
  theme_void()
```

Okay so there are all our centroids, how many do we have in total?

```{r}
nrow(nhgis_centroids)
```

Next, let's pull in the venues data

## Venues

```{r}
venues = read_rds(here("data", "venues_complete.rds"))

glimpse(venues)
```

# Collect travel time data

The goal is to collect travel time data between each venue and each census tract centroid.
How many requests will this be?

```{r}
nrow(nhgis_centroids) * nrow(venues)
```

To do this, we'll be using the Travel [Time Matrix API](https://docs.traveltime.com/api/start/travel-time-distance-matrix) from TravelTime.
According to the documentation, a *request* is made up of up to 10 *searches*.
Each search is either a one-to-many or many-to-one calculation which can include up to 100,000 locations.
So theoretically we could collect all the data we need with a total of 4 requests: 
2 requests of 10 searches (1 search per venue) for each mode of transport, where each search calculates the travel time between a venue and all the tracts.


# Create API Requests

Here is the example of the Time Matrix API request from [TravelTime's documentation](https://docs.traveltime.com/api/reference/distance-matrix):

```{r}
venues |> 
  sample_n(1) |> 
  mutate(lat = st_coordinates(venue_geometry)[,2], 
         lng = st_coordinates(venue_geometry)[,1]) |> 
  st_drop_geometry() |>
  select(id = venue_location, lat, lng) |> 
  pmap(\(id, lat, lng) make_location(id = id, coords = list(lat = as.numeric(lat), lng = as.numeric(lng)))) |> 
  glimpse()
```


```{r}
locationsDF <- data.frame(
  id = c('London center', 'Hyde Park', 'ZSL London Zoo'),
  lat = c(51.508930, 51.508824, 51.536067),
  lng = c(-0.131387, -0.167093, -0.153596)
)
locations <- apply(locationsDF, 1, function(x)
  make_location(id = x['id'], coords = list(lat = as.numeric(x["lat"]),
                                            lng = as.numeric(x["lng"]))))
locations <- unlist(locations, recursive = FALSE)

departure_search <-
  make_search(id = "departure search example",
              departure_location_id = "London center",
              arrival_location_ids = list("Hyde Park", "ZSL London Zoo"),
              departure_time = strftime(as.POSIXlt(Sys.time(), "UTC"), "%Y-%m-%dT%H:%M:%SZ"),
              travel_time = 1800,
              properties = list('travel_time'),
              transportation = list(type = "bus"),
              range = list(enabled = TRUE, width = 600, max_results = 3))

arrival_search <-
  make_search(id = "arrival search example",
              arrival_location_id = "London center",
              departure_location_ids = list("Hyde Park", "ZSL London Zoo"),
              arrival_time = strftime(as.POSIXlt(Sys.time(), "UTC"), "%Y-%m-%dT%H:%M:%SZ"),
              travel_time = 1800,
              properties = list('travel_time', "distance", "distance_breakdown", "fares"),
              transportation = list(type = "public_transport"),
              range = list(enabled = TRUE, width = 600, max_results = 3))

result <-
  time_filter(
    departure_searches = departure_search,
    arrival_searches = arrival_search,
    locations = locations
  )

# print(result)

```

So this gives us a good template of how to create the API requests, but the output isn't the best so I'll make a function similar to that for the isochrones to format the output into a dataframe.

But first, let's organize the searches.
To start, the locations are a dataframe with the id, latitude, and longitude of each location, which they map to the function `make_location()` and then unlist to create a list of 1-row dataframes.

## Locations

```{r}
glimpse(locations)
```

We can do the same with the venues and census tracts

```{r}
# Calculate lat/lon and put them in a list column to match format for API calls
venue_locations_old = venues |>
  mutate(lat = st_coordinates(venue_geometry)[,2],
         lng = st_coordinates(venue_geometry)[,1]) |>
  st_drop_geometry() |>
  select(id = venue_location, lat, lng)

venue_locations_old_list <- apply(venue_locations_old, 1, function(x)
  make_location(id = x['id'], coords = list(lat = as.numeric(x["lat"]),
                                            lng = as.numeric(x["lng"])))) |> 
  unlist(recursive=F)
# locations <- unlist(locations, recursive = FALSE)

venue_locations = venues |> 
  mutate(lat = st_coordinates(venue_geometry)[,2], 
         lng = st_coordinates(venue_geometry)[,1],
         coords = map2(lat, lng, \(lat, lng) list(lat = lat, lng = lng))) |>
  select(id = venue_location, coords) |> 
  st_drop_geometry()

glimpse(venue_locations)
```

```{r}
tract_locations = nhgis_centroids |> 
  mutate(lat = st_coordinates(centroid)[,2], 
         lng = st_coordinates(centroid)[,1]) |> 
  st_drop_geometry() |>
  select(id = TRACTCE, lat, lng) |> 
  pmap(\(id, lat, lng) make_location(id = id, coords = list(lat = as.numeric(lat), lng = as.numeric(lng))))

tract_locations_old = nhgis_centroids |>
  mutate(lat = st_coordinates(centroid)[,2],
         lng = st_coordinates(centroid)[,1]) |>
  st_drop_geometry() |>
  select(id = TRACTCE, lat, lng)

tract_locations_old_list <- apply(tract_locations_old, 1, function(x)
  make_location(id = x['id'], coords = list(lat = as.numeric(x["lat"]),
                                            lng = as.numeric(x["lng"])))) |> 
  unlist(recursive=F)

head(tract_locations) |> glimpse()

```

We can then combine the two dataframes and use `group_split()` to turn them into a list of dataframes, which we pass into the `make_location()` function.

```{r}
locations_list = list_flatten(list(venue_locations, tract_locations))
locations_list_old = list_flatten(list(venue_locations_old_list, tract_locations_old_list))

head(locations_list) |> 
  glimpse()
```


Then to make the departure searches, we need to create a search 

```{r}
glimpse(departure_search)
```

```{r}
tract_ids = map_depth(tract_locations, 2, "id") |> 
  list_simplify()
# venue_ids = map_depth(venue_locations_old, 2, "id") |> 
#   list_simplify()

# departure_search_test = make_search(id = "departure search test",
#                                     departure_location_id = tract_ids,
#                                     arrival_location_ids = venue_ids[[1]],
#                                     departure_time = next_weekday(),
#                                     travel_time = 60*60*3, # 3 hrs
#                                     properties = list('travel_time', 'distance'),
#                                     transportation = list(type = "driving"))
# 
# arrival_search_test <-
#   make_search(id = "arrival search test",
#               arrival_location_id = venue_ids[[1]],
#               departure_location_ids = tract_ids[[1]],
#               arrival_time = next_weekday(),
#               travel_time = 60*60*3,
#               properties = list('travel_time', "distance"),
#               transportation = list(type = "driving"))

# iso_request_df = venues_iso |> 
#   expand_grid(travel_time = 60*c(30, 60, 90, 120),
#               time_of_day = c(9, 11, 13))

time_requests = venue_locations |> sample_n(1) |> 
  expand_grid(tract = map_chr(tract_ids, pluck)) |> 
  # select(id) |> 
  pmap(\(id, coords, tract) make_search(id = id, 
                                        travel_time = 60*60*3, 
                                        arrival_location_id = id,
                                        coords = coords,
                                        departure_location_ids = list(
                                          tract
                                          ),
                                        properties = list('travel_time', 'distance'),
                                        arrival_time = next_weekday(), 
                                        transportation = list(type = "driving")))
```






```{r}

routes(
    # departure_searches = departure_search_test,
    arrival_searches = time_requests[[1]],
    locations = locations_list_old
  )
```

```{r}
time_filter(
    # departure_searches = departure_search_test,
    arrival_searches = time_requests[[1]],
    locations = locations_list_old
  )
```

```{r}
test_df = map(time_requests, \(request) get_routes(request, locations = locations_list_old)) |> 
  list_rbind()
```

```{r}
test_venue = venues |> 
  filter(venue_location == "Universal Studios Lot")

test_df |> 
  mutate(time_mins = travel_time/60) |> 
  right_join(nhgis_la_shp_subset, by = c("id" = "TRACTA")) |> 
  ggplot()+
  geom_sf(aes(geometry = geometry, fill = time_mins))+
  geom_sf(data = test_venue, aes(geometry = venue_geometry), colour = "black", size = 2)+
  scale_fill_viridis_c()+
  coord_sf()+
  theme_void()
```

```{r}
test_df |> 
  mutate(time_mins = travel_time/60) |> 
  ggplot()+
  geom_histogram(aes(x = time_mins), colour = "white")+
  geom_vline(aes(xintercept = mean(time_mins)), colour = "red", linetype = "dashed")+
  geom_text(aes(x = mean(time_mins), y = 200, label = round(mean(time_mins), 1)))+
  labs(title = "Distribution of Travel Times to Universal Studios Lot",
       x = "Travel Time (mins)",
       y = "Density")+
  theme_minimal()
```

Now collect the travel time data for each venue and each tract centroid.

```{r}
time_requests = venue_locations |> #sample_n(1) |> 
  expand_grid(tract = map_chr(tract_ids, pluck)) |> 
  # select(id) |> 
  pmap(\(id, coords, tract) make_search(id = id, 
                                        travel_time = 60*60*3, 
                                        arrival_location_id = id,
                                        coords = coords,
                                        departure_location_ids = list(
                                          tract
                                          ),
                                        properties = list('travel_time', 'distance'),
                                        arrival_time = next_weekday(), 
                                        transportation = list(type = "driving")))
```


```{r}
driving_routes = map(time_requests, \(request) get_routes(request, locations = locations_list_old, sleep_time = 0.5)) |> 
  list_rbind()
```


```{r}
# Create a combination of venues and travel times for isochrone requests
travel_times = 60*c(30, 60, 90, 120)

iso_request_df = venues_iso |> 
  expand_grid(travel_time = 60*c(30, 60, 90, 120),
              time_of_day = c(9, 11, 13))

glimpse(iso_request_df)
```



### Transit

```{r}
# Create a search object for each venue: public transport
venues_iso_search = iso_request_df |> 
  pmap(\(id, coords, travel_time, time_of_day) make_search(id = str_c(id, "_", travel_time), 
                                              travel_time = travel_time, 
                                              arrival_time = next_weekday(time_in_hrs = time_of_day),#"2024-12-18T09:00:00-08:00", 
                                              coords = coords, 
                                              transportation = list(type = "public_transport", 
                                                                    walking_time=15*60, 
                                                                    pt_change_delay = 120),
                                              level_of_detail = list(scale_type = "simple", level = "medium")))
```


