---
title: "Collecting Tract-level Travel Time Data"
author: "Mark Barrett"
format: html
---

Documentation:

- [Github](https://github.com/traveltime-dev/traveltime-sdk-r)
- [Routes](https://playground.traveltime.com/routes)

# Plan

Analyze the differences in travel time to Los Angeles 2028 Olympic venues between driving and transit.
Collect tract-level travel time data to each venue using both transit and driving modes.
To do this, I will find the centroid of each tract in the Los Angeles area and then use the TravelTime API to collect travel time data to each venue.

- 30 mins, 1hr, and 1.5hrs
- via car and via public transport
- at 8am and 11am on a Weekday

```{r setup}
library(here)
library(tidyverse)
library(sf)
library(traveltimeR)
library(jsonlite)

# Load custom functions
source(here("R", "fct-traveltime.R"))
```

# Data

## Census Tracts

Downloaded from nhgis.com

```{r}
# Census tract shapefiles
nhgis_shp_file = here("data", "nhgis", "US_tract_2020.shp")

# Create list of tracts to exclude
exclude_tracts = tribble(
  ~TRACTCE, ~include,
  "599100", F, # Catalina Island
  "599000", F, # Catalina Island
  "930400", F,  # Angeles Forest
)

# Load the shapefile
nhgis_shp = read_sf(nhgis_shp_file) |> 
  # Filter to Los Angeles, CA
  filter(STATEFP == "06", COUNTYFP == "037") |>
  # Reproject CRS to WGS84
  st_transform(crs = "EPSG:4326") |> 
  # Create an opinionated subset of LA tracts
  # anti_join(exclude_tracts, by = "TRACTCE") |>
  # Remove tracts where the centroid is north of 34.35
  # filter(st_coordinates(st_centroid(geometry))[,2] < 34.35) |>
  select(GISJOIN, TRACTCE, GEOID, NAME, NAMELSAD, geometry) |> 
  left_join(exclude_tracts, by = "TRACTCE") |> 
  mutate(include = case_when(!is.na(include) ~ include,
                             st_coordinates(st_centroid(geometry))[,2] > 34.35 ~ F,
                             .default = T))

glimpse(nhgis_shp)

```

Check how this looks on a map

```{r}
ggplot(nhgis_shp) +
  geom_sf(aes(fill = include), colour=NA) +
  coord_sf() +
  labs(title = "Census Tracts in Los Angeles County",
       subtitle = "Filtered to exclude Catalina Island, Angeles Forest, and Antelope Valley")+
  theme_void()
```

Remove these tracts

```{r}
nhgis_la = nhgis_shp |> 
  filter(include) |> 
  select(-include)
```


# Find centroids of census tracts

```{r}
nhgis_centroids = nhgis_la |> 
  # Find centroids of each tract
  mutate(centroid = st_centroid(geometry))

ggplot(nhgis_centroids) +
  geom_sf(aes(geometry = geometry), fill = NA) +
  geom_sf(aes(geometry = centroid), colour = "red")+
  coord_sf() +
  labs(title = "Census Tracts centroids")+
  theme_void()
```

Okay so there are all our centroids, how many do we have in total?

```{r}
nrow(nhgis_centroids)
```

Next, let's pull in the venues data

## Venues

```{r}
venues = read_rds(here("data", "venues_complete.rds"))

glimpse(venues)
```

# Collect travel time data

The goal is to collect travel time data between each venue and each census tract centroid.
How many requests will this be?

```{r}
nrow(nhgis_centroids) * nrow(venues)
```

To do this, we'll be using the Travel [Time Matrix API](https://docs.traveltime.com/api/start/travel-time-distance-matrix) from TravelTime.
According to the documentation, a *request* is made up of up to 10 *searches*.
Each search is either a one-to-many or many-to-one calculation which can include up to 100,000 locations.
So theoretically we could collect all the data we need with a total of 4 requests: 
2 requests of 10 searches (1 search per venue) for each mode of transport, where each search calculates the travel time between a venue and all the tracts.


# Create API Requests

Here is the example of the Time Matrix API request from [TravelTime's documentation](https://docs.traveltime.com/api/reference/distance-matrix):

```{r}
venues |> 
  sample_n(1) |> 
  mutate(lat = st_coordinates(venue_geometry)[,2], 
         lng = st_coordinates(venue_geometry)[,1]) |> 
  st_drop_geometry() |>
  select(id = venue_location, lat, lng) |> 
  pmap(\(id, lat, lng) make_location(id = id, coords = list(lat = as.numeric(lat), lng = as.numeric(lng)))) |> 
  glimpse()
```


```{r}
locationsDF <- data.frame(
  id = c('London center', 'Hyde Park', 'ZSL London Zoo'),
  lat = c(51.508930, 51.508824, 51.536067),
  lng = c(-0.131387, -0.167093, -0.153596)
)
locations <- apply(locationsDF, 1, function(x)
  make_location(id = x['id'], coords = list(lat = as.numeric(x["lat"]),
                                            lng = as.numeric(x["lng"]))))
locations <- unlist(locations, recursive = FALSE)

departure_search <-
  make_search(id = "departure search example",
              departure_location_id = "London center",
              arrival_location_ids = list("Hyde Park", "ZSL London Zoo"),
              departure_time = strftime(as.POSIXlt(Sys.time(), "UTC"), "%Y-%m-%dT%H:%M:%SZ"),
              travel_time = 1800,
              properties = list('travel_time'),
              transportation = list(type = "bus"),
              range = list(enabled = TRUE, width = 600, max_results = 3))

arrival_search <-
  make_search(id = "arrival search example",
              arrival_location_id = "London center",
              departure_location_ids = list("Hyde Park", "ZSL London Zoo"),
              arrival_time = strftime(as.POSIXlt(Sys.time(), "UTC"), "%Y-%m-%dT%H:%M:%SZ"),
              travel_time = 1800,
              properties = list('travel_time', "distance", "distance_breakdown", "fares"),
              transportation = list(type = "public_transport"),
              range = list(enabled = TRUE, width = 600, max_results = 3))

result <-
  time_filter(
    departure_searches = departure_search,
    arrival_searches = arrival_search,
    locations = locations
  )

# print(result)

```

So this gives us a good template of how to create the API requests, but the output isn't the best so I'll make a function similar to that for the isochrones to format the output into a dataframe.

But first, let's organize the searches.
To start, the locations are a dataframe with the id, latitude, and longitude of each location, which they map to the function `make_location()` and then unlist to create a list of 1-row dataframes.

## Locations

```{r}
glimpse(locations)
```

We can do the same with the venues and census tracts

```{r}
# Calculate lat/lon and put them in a list column to match format for API calls
venue_locations_old = venues |>
  mutate(lat = st_coordinates(venue_geometry)[,2],
         lng = st_coordinates(venue_geometry)[,1]) |>
  st_drop_geometry() |>
  select(id = venue_location, lat, lng)

venue_locations_old_list <- apply(venue_locations_old, 1, function(x)
  make_location(id = x['id'], coords = list(lat = as.numeric(x["lat"]),
                                            lng = as.numeric(x["lng"])))) |> 
  unlist(recursive=F)
# locations <- unlist(locations, recursive = FALSE)

venue_locations = venues |> 
  mutate(lat = st_coordinates(venue_geometry)[,2], 
         lng = st_coordinates(venue_geometry)[,1],
         coords = map2(lat, lng, \(lat, lng) list(lat = lat, lng = lng))) |>
  select(id = venue_location, coords) |> 
  st_drop_geometry()

glimpse(venue_locations)
```

```{r}
tract_locations = nhgis_centroids |> 
  mutate(lat = st_coordinates(centroid)[,2], 
         lng = st_coordinates(centroid)[,1]) |> 
  st_drop_geometry() |>
  select(id = TRACTCE, lat, lng) |> 
  pmap(\(id, lat, lng) make_location(id = id, coords = list(lat = as.numeric(lat), lng = as.numeric(lng))))

tract_locations_old = nhgis_centroids |>
  mutate(lat = st_coordinates(centroid)[,2],
         lng = st_coordinates(centroid)[,1]) |>
  st_drop_geometry() |>
  select(id = TRACTCE, lat, lng)

tract_locations_old_list <- apply(tract_locations_old, 1, function(x)
  make_location(id = x['id'], coords = list(lat = as.numeric(x["lat"]),
                                            lng = as.numeric(x["lng"])))) |> 
  unlist(recursive=F)

head(tract_locations) |> glimpse()

```

We can then combine the two dataframes and use `group_split()` to turn them into a list of dataframes, which we pass into the `make_location()` function.

```{r}
locations_list = list_flatten(list(venue_locations, tract_locations))
locations_list_old = list_flatten(list(venue_locations_old_list, tract_locations_old_list))

head(locations_list) |> 
  glimpse()
```


Then to make the departure searches, we need to create a search 

```{r}
glimpse(departure_search)
```

```{r}
tract_ids = map_depth(tract_locations, 2, "id") |> 
  list_simplify()
# venue_ids = map_depth(venue_locations_old, 2, "id") |> 
#   list_simplify()

# departure_search_test = make_search(id = "departure search test",
#                                     departure_location_id = tract_ids,
#                                     arrival_location_ids = venue_ids[[1]],
#                                     departure_time = next_weekday(),
#                                     travel_time = 60*60*3, # 3 hrs
#                                     properties = list('travel_time', 'distance'),
#                                     transportation = list(type = "driving"))
# 
# arrival_search_test <-
#   make_search(id = "arrival search test",
#               arrival_location_id = venue_ids[[1]],
#               departure_location_ids = tract_ids[[1]],
#               arrival_time = next_weekday(),
#               travel_time = 60*60*3,
#               properties = list('travel_time', "distance"),
#               transportation = list(type = "driving"))

# iso_request_df = venues_iso |> 
#   expand_grid(travel_time = 60*c(30, 60, 90, 120),
#               time_of_day = c(9, 11, 13))

time_requests = venue_locations |> sample_n(1) |> 
  expand_grid(tract = map_chr(tract_ids, pluck)) |> 
  # select(id) |> 
  pmap(\(id, coords, tract) make_search(id = id, 
                                        travel_time = 60*60*3, 
                                        arrival_location_id = id,
                                        coords = coords,
                                        departure_location_ids = list(
                                          tract
                                          ),
                                        properties = list('travel_time', 'distance'),
                                        arrival_time = next_weekday(), 
                                        transportation = list(type = "driving")))
```






```{r}

routes(
    # departure_searches = departure_search_test,
    arrival_searches = time_requests[[1]],
    locations = locations_list_old
  )
```

```{r}
time_filter(
    # departure_searches = departure_search_test,
    arrival_searches = time_requests[[1]],
    locations = locations_list_old
  )
```

```{r}
test_df = map(time_requests, \(request) get_routes(request, locations = locations_list_old)) |> 
  list_rbind()
```

```{r}
test_venue = venues |> 
  filter(venue_location == "Universal Studios Lot")

test_df |> 
  mutate(time_mins = travel_time/60) |> 
  right_join(nhgis_la_shp_subset, by = c("id" = "TRACTA")) |> 
  ggplot()+
  geom_sf(aes(geometry = geometry, fill = time_mins))+
  geom_sf(data = test_venue, aes(geometry = venue_geometry), colour = "black", size = 2)+
  scale_fill_viridis_c()+
  coord_sf()+
  theme_void()
```

```{r}
test_df |> 
  mutate(time_mins = travel_time/60) |> 
  ggplot()+
  geom_histogram(aes(x = time_mins), colour = "white")+
  geom_vline(aes(xintercept = mean(time_mins)), colour = "red", linetype = "dashed")+
  geom_text(aes(x = mean(time_mins), y = 200, label = round(mean(time_mins), 1)))+
  labs(title = "Distribution of Travel Times to Universal Studios Lot",
       x = "Travel Time (mins)",
       y = "Density")+
  theme_minimal()
```

Now collect the travel time data for each venue and each tract centroid.

```{r}
time_requests = venue_locations |> #sample_n(1) |> 
  filter(id == "Sepulveda Basin Recreation Area") |> 
  expand_grid(tract = map_chr(tract_ids, pluck)) |> 
  # select(id) |> 
  pmap(\(id, coords, tract) make_search(id = str_c(id, " | ", tract), 
                                        travel_time = 60*60*3, 
                                        arrival_location_id = id,
                                        coords = coords,
                                        departure_location_ids = list(
                                          tract
                                          ),
                                        properties = list('travel_time', 'distance'),
                                        arrival_time = next_weekday(), 
                                        transportation = list(type = "driving")))
```

## Test out Time Filter (Fast) 

```{r}
arrival_search_test = make_search(id = "time filter test",
            departure_location_ids = tract_ids,
            arrival_location_id = venue_locations_old_list[[1]]$id,
              arrival_time_period = "weekday_morning",
            # arrival_time = next_weekday(),
            travel_time = 60*60*3, # 3 hrs
            properties = list("travel_time"),
            transportation = list(type = "driving"))

result = time_filter_fast(
  arrival_many_to_one = arrival_search_test,
  locations = locations_list_old
)

result_df = result$contentParsed$results[[1]]$locations |> 
    map_depth(1, map_if, is_list, list_flatten) |> 
    map_if(is_list, list_flatten, name_spec = "{inner}") |> 
    map(as_tibble) |> 
    list_rbind()
```

```{r}
time_requests = venue_locations |> 
  select(-coords) |> 
  expand_grid(transportation_type = c("driving", "public_transport")) |>
  pmap(\(id, transportation_type) make_search(id = str_c(id, " | ", transportation_type), 
                         departure_location_ids = tract_ids,
                         arrival_location_id = id,
                         arrival_time_period = "weekday_morning",
                         travel_time = 60*60*3, # 3 hrs
                         properties = list("travel_time", "distance"),
                         transportation = list(type = transportation_type, 
                                               walking_time=15*60, 
                                               pt_change_delay = 120)))

# time_requests[[1]]

# driving_times = map(time_requests, \(request) time_filter_fast_to_df(search = request, locations = locations_list_old))

# list_rbind(driving_times)

travel_times = map(time_requests, \(request) time_filter_fast_to_df(search = request, locations = locations_list_old, sleep_time = 0))

list_rbind(travel_times) |> 
  mutate(time_mins = travel_time/60,
         distance_mi = distance/1609, .before = "travel_time") |> 
  arrange(id, venue_id, transportation) |> 
  View("Time Filter Fast")
```

Test out time filter

```{r}
travel_times2 = time_filter_to_df(venue_locations, tract_ids, locations_list_old, 0)

# requests = venue_locations |> 
#     select(-coords) |> 
#     expand_grid(transportation_type = c("driving", "public_transport"), 
#                 tract_list = list(tract_ids[1:length(tract_ids)/2], tract_ids[(length(tract_ids)/2)+1:length(tract_ids)])) |>
#     pmap(\(id, transportation_type, tract_list) make_search(id = str_c(id, " | ", transportation_type), 
#                                                             departure_location_ids = tract_list,
#                                                             arrival_location_id = id,
#                                                             arrival_time = next_weekday(),
#                                                             travel_time = 60*60*3, # 3 hrs
#                                                             properties = list("travel_time", "distance"),
#                                                             transportation = list(type = transportation_type, 
#                                                                                   walking_time=15*60, 
#                                                                                   pt_change_delay = 120)))
# 
# map(requests, \(request) time_filter_fct(request, tract_ids, locations_list_old)) |> 
#     list_rbind()
# 
# requests[[2]]

travel_times2 |> 
  # mutate(time_mins = travel_time/60,
  #        distance_mi = distance/1609, .before = "travel_time") |> 
  arrange(id, venue_id, transportation) |> 
  View("Time Filter")
```

```{r}
saveRDS(travel_times2, here("data", str_c("travel_times_", today(),".rds")))
```

```{r}
travel_times2 |> #glimpse()
  filter(str_detect(venue_id, "Sepulveda Basin Recreation Area")) |>
  group_by(transportation) |> 
  mutate(time_mins = if_else(transportation == "driving", time_mins*1.2, time_mins),
         mean_time = mean(time_mins)) |> 
  ungroup() |> 
  filter(time_mins < 400) |> 
  ggplot()+
  geom_histogram(aes(x = time_mins, fill = transportation, y = stat(width*density)), 
                 binwidth = 5,
                 colour = "white")+
  geom_vline(aes(xintercept = mean_time, colour = transportation), linetype = "dashed")+
  scale_x_continuous(breaks = seq(0, 240, 30))+
  scale_y_continuous(breaks = scales::breaks_pretty(), labels = scales::percent)+
  # scale_alpha_discrete(breaks = c("driving", "public_transport"), range = c(1, .025))+
  # geom_text(aes(x = mean_time, y = .3, label = round(mean_time, 1)))+
  labs(title = "Distribution of Travel Times to Sepulveda Basin Recreation Area",
       x = "Travel Time (mins)")+
  # facet_grid(rows = vars(transportation), scales = "free_y", axes = "all_x",
  #            switch = "y")+
  theme_minimal()+
  theme()
```



Compare results to Google Maps estimates

```{r}
results_diff = tribble(
  ~from_tract, ~to_venue,                       ~transportation, ~method,   ~time_mins,
  "101110",    "Arena Inglewood (Intuit Dome)", "driving",       "fast",    40,
  "101110",    "Arena Inglewood (Intuit Dome)", "driving",       "normal",  39.8,
  "101110",    "Arena Inglewood (Intuit Dome)", "driving",       "google",  75,
  "101110",    "Arena Inglewood (Intuit Dome)", "transit",       "fast",    112.4,
  "101110",    "Arena Inglewood (Intuit Dome)", "transit",       "normal",  144.9,
  "101110",    "Arena Inglewood (Intuit Dome)", "transit",       "google",  135,
  "101110",    "LA Live", "driving",       "fast",    24.3,
  "101110",    "LA Live", "driving",       "normal",  24.9,
  "101110",    "LA Live", "driving",       "google",  35,
  "101110",    "LA Live", "transit",       "fast",    73.6,
  "101110",    "LA Live", "transit",       "normal",  98.9,
  "101110",    "LA Live", "transit",       "google",  97,
  "190102",    "Dodger Stadium", "driving",       "fast",    9.1,
  "190102",    "Dodger Stadium", "driving",       "normal",    17.1,
  "190102",    "Dodger Stadium", "driving",       "google",    35,
  "190102",    "Dodger Stadium", "transit",       "fast",    46.2,
  "190102",    "Dodger Stadium", "transit",       "normal",    54.9,
  "190102",    "Dodger Stadium", "transit",       "google",    61,
  "199400",    "Arena Inglewood (Intuit Dome)", "driving",       "fast",    24.5,
  "199400",    "Arena Inglewood (Intuit Dome)", "driving",       "normal",  25,
  "199400",    "Arena Inglewood (Intuit Dome)", "driving",       "google",  45,
  "199400",    "Arena Inglewood (Intuit Dome)", "transit",       "fast",    83.3,
  "199400",    "Arena Inglewood (Intuit Dome)", "transit",       "normal",  86.8,
  "199400",    "Arena Inglewood (Intuit Dome)", "transit",       "google",  88,
  "199400",    "Belmont Shore", "driving",       "fast",  39.3,
  "199400",    "Belmont Shore", "driving",       "normal", 40.7,
  "199400",    "Belmont Shore", "driving",       "google", 50,
  "199400",    "Belmont Shore", "transit",       "fast", 122.4,
  "199400",    "Belmont Shore", "transit",       "normal", 111,
  "199400",    "Belmont Shore", "transit",       "google", 112
  
  
)
```


```{r}
driving_routes = map(time_requests, \(request) get_routes(request, locations = locations_list_old, sleep_time = .8)) |> 
  list_rbind()
```


```{r}
driving_routes |> 
  mutate(time_mins = travel_time/60) |> 
  ggplot()+
  geom_histogram(aes(x = time_mins), colour = "white")+
  geom_vline(aes(xintercept = mean(time_mins)), colour = "red", linetype = "dashed")+
  geom_text(aes(x = mean(time_mins), y = 200, label = round(mean(time_mins), 1)))+
  labs(title = "Distribution of Travel Times to Sepulveda Basin Recreation Area",
       x = "Travel Time (mins)",
       y = "Density")+
  theme_minimal()
```



### Transit

```{r}
# Create a search object for each venue: public transport
transit_time_requests = venue_locations |> #sample_n(1) |> 
  filter(id == "Sepulveda Basin Recreation Area") |> 
  expand_grid(tract = map_chr(tract_ids, pluck)) |> 
  # select(id) |> 
  pmap(\(id, coords, tract) make_search(id = str_c(id, " | ", tract), 
                                        travel_time = 60*60*3, 
                                        arrival_location_id = id,
                                        coords = coords,
                                        departure_location_ids = list(
                                          tract
                                          ),
                                        properties = list('travel_time', 'distance'),
                                        arrival_time = next_weekday(), 
                                        transportation = list(type = "public_transport", 
                                                                    walking_time=15*60, 
                                                                    pt_change_delay = 120)))
```

```{r}
transit_routes = map(transit_time_requests, \(request) get_routes(request, locations = locations_list_old, sleep_time = .8)) |> 
  list_rbind()
```

```{r}
transit_routes |> 
  mutate(time_mins = travel_time/60) |> 
  ggplot()+
  geom_histogram(aes(x = time_mins), colour = "white")+
  geom_vline(aes(xintercept = mean(time_mins)), colour = "red", linetype = "dashed")+
  geom_text(aes(x = mean(time_mins), y = 200, label = round(mean(time_mins), 1)))+
  labs(title = "Distribution of Transit Times to Sepulveda Basin Recreation Area",
       x = "Travel Time (mins)")+
  theme_minimal()
```


```{r}
routes_combined = list_rbind(list(driving_routes, transit_routes))

routes_combined |> 
  group_by(transportation) |> 
  mutate(time_mins = travel_time/60,
         mean_time = mean(time_mins)) |> 
  ungroup() |> 
  filter(time_mins < 400) |> 
  ggplot()+
  geom_histogram(aes(x = time_mins, fill = transportation, y = stat(width*density)), 
                 binwidth = 5,
                 colour = "white")+
  geom_vline(aes(xintercept = mean_time, fill = transportation), colour = "red", linetype = "dashed")+
  scale_x_continuous(breaks = seq(0, 240, 30))+
  scale_y_continuous(breaks = scales::breaks_pretty(), labels = scales::percent)+
  # geom_text(aes(x = mean_time, y = .3, label = round(mean_time, 1)))+
  labs(title = "Distribution of Travel Times to Sepulveda Basin Recreation Area",
       x = "Travel Time (mins)")+
  facet_grid(rows = vars(transportation), scales = "free_y", axes = "all_x",
             switch = "y")+
  theme_minimal()+
  theme()
```

```{r}
routes_combined |> 
  glimpse()
```

Gather all the requests in one object

```{r}
driving_time_requests = venue_locations |> #sample_n(1) |> 
  expand_grid(tract = map_chr(tract_ids, pluck)) |> 
  # select(id) |> 
  pmap(\(id, coords, tract) make_search(id = str_c(id, " | ", tract, " | driving"), 
                                        travel_time = 60*60*3, 
                                        arrival_location_id = id,
                                        coords = coords,
                                        departure_location_ids = list(
                                          tract
                                          ),
                                        properties = list('travel_time', 'distance'),
                                        arrival_time = next_weekday(), 
                                        transportation = list(type = "driving")))

# Create a search object for each venue: public transport
transit_time_requests = venue_locations |> #sample_n(1) |> 
  expand_grid(tract = map_chr(tract_ids, pluck)) |> 
  # select(id) |> 
  pmap(\(id, coords, tract) make_search(id = str_c(id, " | ", tract, " | transit"), 
                                        travel_time = 60*60*3, 
                                        arrival_location_id = id,
                                        coords = coords,
                                        departure_location_ids = list(
                                          tract
                                          ),
                                        properties = list('travel_time', 'distance'),
                                        arrival_time = next_weekday(), 
                                        transportation = list(type = "public_transport", 
                                                                    walking_time=15*60, 
                                                                    pt_change_delay = 120)))

```

```{r}
driving_routes = map(driving_time_requests, \(request) get_routes(request, locations = locations_list_old, sleep_time = 1), .progress = T) |> 
  list_rbind()

transit_routes = map(transit_time_requests, \(request) get_routes(request, locations = locations_list_old, sleep_time = .8)) |> 
  list_rbind()
```



